---
title: "CMIP6_carbon"
output: html_notebook
---

A look at the carbon cycle in CMIP6.


Questions for Andy: How did he choose these models? Which outputs are a priority?

Notes:
One aim is to build some capacity for reading CMIP6 data.

Understanding CMIP5/6 filenames 

http://cerfacs.fr/~coquart/data/uploads/cmip6_global_attributes_filenames_cvs_v6.2.6.pdf

http://clipc-services.ceda.ac.uk/dreq/index/miptable.html


"gn" vs "gr" vs "gr1"
https://github.com/WCRP-CMIP/CMIP6_CVs/blob/master/CMIP6_grid_label.json



```{r}
library(ncdf4)

source('explore-JULES-ES-1p0_PPE_functions.R')
```

```{r}
extractModelNames = function(filenames, startstring, stopstring){
  # Function to extract model names from a vector of cmip filenames
  # that have similar startstring and stopstring
  
  startstring_length <- nchar(startstring)
  
  fs <- lapply(filenames,regexpr, pattern = startstring)
  fe <- lapply(filenames,regexpr, pattern = stopstring)
  out <- rep(NA, length(filenames))
  for(i in 1:length(filenames)){
    out[i] = substr(filenames[i], attr(fs[i][[1]], 'match.length')+1+startstring_length, fe[[i]][1]-2)
  }
  out
}
```


```{r}

extractCmip6Timeseries <- function(datadir, filelist, var, nyr){
  
  
  out <- matrix(nrow = length(filelist), ncol = nyr)
  
  for(i in 1:length(filelist)){
    # Open and extract data from the netcdf file
    fn <- filelist[i]
    fnpath <- paste0(datadir, fn)
    #print(fnpath)
    nc <- nc_open(fnpath)
    v <- ncvar_get(nc, var)
    #print(ncatt_get(nc, 'time')$units)
    nc_close(nc)
    
    vtail <- tail(v, nyr)
    vtail_len <- length(vtail)
    #print(length(vtail))
    out[i, ((nyr - vtail_len)+1):nyr ] <- vtail
  }
  
  out
  
}
```



```{r}
hyrs = 1850:2014
nyr = length(hyrs)

datadir <- '/scratch/hadaw/temp/ForDoug/'

linecol <- makeTransparent('black', 100)
```

## NBP and cumulative land flux

Land flux calculated here as the cumulative sum of nbp

```{r}

# start with nbp
filelist_nbp <- dir(datadir, pattern = 'nbp')
nbp <- extractCmip6Timeseries(datadir = datadir, filelist = filelist_nbp, var = 'nbp', nyr = nyr)

```


```{r}

nbp_scaled <- nbp * 86400*365 /1e12

land_flux <- apply(nbp_scaled, 2, cumsum)
matplot(t(land_flux), type = 'l', ylim = c(-200, 200), col = linecol,
        main = 'net land flux', xlab = 'year', ylab = 'net land flux (GtC)')


```


## Vegetation and soil carbon

We should compare CMIP6 carbon cycle with

https://dougmcneall.github.io/jules_ppe/constrain-JULES-ES-1p0.nb.html#14_Plot_timeseries_of_aggregated_global_timeseries

```{r}

# Soil carbon
filelist_cSoil <- dir(datadir, pattern = 'cSoil')
cSoil <- extractCmip6Timeseries(datadir = datadir, filelist = filelist_cSoil, var = 'cSoil', nyr = nyr)
cSoil_scaled <- cSoil /1e12
cSoil_scaled_anom <- anomalizeTSmatrix(cSoil_scaled, ix = 1:2)


# Vegetation carbon
filelist_cVeg <- dir(datadir, pattern = 'cVeg')
cVeg <- extractCmip6Timeseries(datadir = datadir, filelist = filelist_cVeg, var = 'cVeg', nyr = nyr)
cVeg_scaled <- cVeg /1e12
cVeg_scaled_anom <- anomalizeTSmatrix(cVeg_scaled, ix = 1:2)

```


Global soil carbon varies in the ensemble between 0 and 2500 Gt, although it would be good to understand if there are some processing errors in Andy's code or my interpretation of it.

The anomaly makes more sense, varying from around -100 to +200 GtC by the modern day.


```{r, fig.width = 8, fig.height = 8}

# Plot soil carbon
par(mfrow = c(1,2))
matplot(hyrs, t(cSoil_scaled), type = 'l', col = linecol, lty = 'solid',
        xlab = 'year', main = 'soil carbon', ylab = 'soil carbon (GtC)')

matplot(hyrs, t(cSoil_scaled_anom), type = 'l', col = linecol, lty = 'solid',
        xlab = 'year', main = 'anomaly', ylab = 'soil carbon anomaly (GtC)')

```

Global vegetation carbon varies in the ensemble between 100 and 6000 Gt, although the vast majority are below 750 GtC at present day.

The anomaly varies from around -100 to +100 GtC by the modern day.

```{r, fig.width = 8, fig.height = 8}

# Plot vegetation carbon
par(mfrow = c(1,2))
matplot(hyrs, t(cVeg_scaled), type = 'l', col = linecol, lty = 'solid',
        xlab = 'years', ylab = 'vegetation carbon (GtC)', main = 'vegetation carbon' 
        )

matplot(hyrs, t(cVeg_scaled_anom), type = 'l', col = linecol, lty = 'solid', 'vegetation carbon (GtC)', main = 'anomaly')

```


