---
title: "Compare the carbon cycle hen JULES driven by reanalysis and UKESM"
output: html_notebook
---

Compare historic JULES when run driven by reanalysis and when driven by UKESM.

Things to check:

I think the 500 members of S3 are wave01 inputs. That would make sense of the 500 members plus 501st member being the standard.

```{r}

# running in /net/home/h01/hadda/jules_ppe/docs
#global_mean_timeseries_wave00_file <- "constrain_JULES_ES_1p0/data/global_mean_timeseries_wave00_2022-09-13.rdata"

#load(global_mean_timeseries_wave00_file)


source("JULES-ES-1p0-common-packages.R")
source("JULES-ES-1p0-common-functions.R")
```

```{r}
historical_carbon_budget <- read_excel('Global_Carbon_Budget_2020v1.0.xlsx', sheet = "Historical Budget", skip = 15, n_max = 270)
match_years_ix <- which(historical_carbon_budget$Year %in% 1850:2013)

land_sink_net <- historical_carbon_budget$`land sink` - historical_carbon_budget$`land-use change emissions`

match_years <- historical_carbon_budget$Year[match_years_ix]
cumulative_net_land_sink <- cumsum(land_sink_net[match_years_ix])


par(las = 1)
plot(match_years, cumulative_net_land_sink,
     type = 'l', main = "Cumulative Net Land Sink", xlab = "", ylab = "GtC",
     ylim = c(-80, 20))

```



```{r}


global_mean_timeseries_stan_file <- "constrain_JULES_ES_1p0/data/global_mean_timeseries_stan_2022-09-13.rdata"

load(global_mean_timeseries_stan_file)

```


Modern value wave 01
```{r}
#ensemble_wave01_file <- "constrain_JULES_ES_1p0/data/ensemble_wave01_2022-04-08.rdata"
  
#load(ensemble_wave01_file)


```

Full timeseries wave01
```{r}

ensemble_wave01_ts_file <-  "constrain_JULES_ES_1p0/data/ensemble_timeseries_wave01_2022-09-14.rdata"

load(ensemble_wave01_ts_file)


```


# S3 (wave01 locations)
```{r}

load("ensemble_timeseries_ssp_2022-08-09.rdata")

```

```{r}

plot(npp_ens_wave01[,1],npp_ens_ssp585_S3[1:400,1] )
abline(0,1)


```

```{r}


plot(npp_ens_wave01[,164],npp_ens_ssp585_S3[1:400,164] )
abline(0,1)

```



```{r, fig.width = 7, fig.height = 9}

# Plot cumulative NBP

cnbp_ens_wave01 <- t(apply(nbp_ens_wave01, 1, FUN = cumsum))
cnbp_ens_ssp585 <- t(apply(nbp_ens_ssp585_S3, 1, FUN = cumsum))

cnbp_stan_wave01 <- cumsum(nbp_stan)
cnbp_stan_ssp585 <- cnbp_ens_ssp585[501, ]

wave01_years <- 1850:2013
ssp585_years <- 1850:2100

par(las = 1)
matplot(wave01_years, t(cnbp_ens_wave01), type = 'l', lty = 'solid', col = makeTransparent('tomato3', 120), xlab = '',
        bty = 'n', 
        ylab = 'Cumulative NBP (GtC)', 
        ylim = c(-100,300), lwd = 2)

 matlines(ssp585_years, t(cnbp_ens_ssp585) , col = makeTransparent('skyblue3', 120), lty = 'solid', lwd = 2)

lines(wave01_years, cnbp_stan_wave01, col = 'red', lty = 'dashed', lwd = 2)
lines(ssp585_years, cnbp_stan_ssp585, col = 'blue', lty = 'dashed', lwd = 2)
lines(match_years, cumulative_net_land_sink, col = 'black', lty = 'dashed', lwd = 2)
legend('topleft', 
       legend = c('Observation-forced', 'Model-forced', 'Standard member'),
       col = c(makeTransparent('tomato3', 120), makeTransparent('skyblue3', 120), 'darkgrey'), 
       lty = c('solid', 'solid', 'dashed' ),
       lwd = 2)


```


Cumulative NBP driven by UKESM is significantly lower than driven by reanalysis.
```{r, fig.width = 7, fig.height = 7}


plot(cnbp_ens_wave01[1:400, 164], cnbp_ens_ssp585[1:400, 164],
     xlim = c(-100, 300), ylim = c(-100, 300),
     main = "Cumulative NBP 1850-2014",
     xlab = "Observations-forced",
    ylab = "Model-forced")
abline(0,1)



```
On average, CNBP is 25 GtC higher in obs forced runs than model forced runs
```{r, fig.width = 7, fig.height = 7}


cnbp_ens_wave01_outliers_ix <-  which(cnbp_ens_wave01[1:400, 164] < -1000 | cnbp_ens_wave01[1:400, 164] > 1000)
cnbp_ens_ssp585_outliers_ix <- which(cnbp_ens_ssp585[1:400, 164] < -1000 | cnbp_ens_ssp585[1:400, 164] > 1000) # none



cnbp_wave01 <- cnbp_ens_wave01[(1:400)[-cnbp_ens_wave01_outliers_ix], 164]
cnbp_ssp585 <- cnbp_ens_ssp585[(1:400)[-cnbp_ens_wave01_outliers_ix], 164]

fit <- lm(cnbp_wave01 ~ cnbp_ssp585)

# coef(fit)[2] *100
# ssp585 is 77% of reanalysis driven cnbp

plot(cnbp_ssp585,cnbp_wave01,
          main = "Cumulative NBP 1850-2014",
     ylab = "Observations-forced",
    xlab = "Model-forced")
abline(0,1)
abline(coef(fit), col = 'red')
legend('topleft', legend = c('x=y', 'fit'), lty = c('solid', 'solid'), col = c('black', 'red'))




```
```{r}


hist(cnbp_wave01 - cnbp_ssp585, main = "Obs forced - model forced CNBP")

mdcnbp <- mean(cnbp_wave01 - cnbp_ssp585)

text(100, 100, paste0('Mean difference = ',round(mdcnbp,1)))
abline(v = mdcnbp, col = 'grey', lty = 'dashed')



```


