---
title: "Analysis of JULES1p0 ensemble SSP585"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 2
    number_sections: yes
---


```{r, echo = FALSE, message = FALSE, warning=FALSE, results = 'hide'}
# Load helper functions

knitr::opts_chunk$set(fig.path = "figs/", echo = FALSE, message = FALSE, warnings = FALSE)


```

```{r}

source("JULES-ES-1p0-common-packages.R")
source("JULES-ES-1p0-common-functions.R")
```


```{r}

linePlotMultiEns <- function(years, ens1, ens2, col1, col2, ylab, main, ylim = NULL){
  # Plot wave00 and wave01 timeseries on top of one another
  
  nt <- length(years) 
  if(is.null(ylim)){
    
  #ylim = range(c(ens1[,1], ens1[,nt], ens2[,1], ens2[ ,nt]))
  ylim = range(c(ens1,ens2))
  }
  
  else ylim <- ylim
  
  matplot(years, t(ens1), type = 'l', lty = 'solid',ylim = ylim, col = col1,
        ylab = ylab, main = main, xlab = '',
        bty = 'n', lwd = 1.5)
  matlines(years, t(ens2), col = col2, lty = 'solid', lwd = 1.5)
}


makeTimeseriesEnsembleSSP <- function(ensloc, variable, nstart, nend, cn = 1850:2100){
  
  ysec <- 31536000
  
  nens <- (nend - nstart) + 1
  # nens is number of ensemble members
  # nts length of timeseries
  # cn is colnames()
  datmat <- matrix(NA, nrow = nens, ncol = length(cn))
  colnames(datmat) <- cn
  
  enslist <- paste("P", formatC(nstart:nend, width=4, flag="0"), sep="")
  
  for(i in 1:nens){
    
    ensmember <- enslist[i]
    
    fn <- paste0(ensloc,'JULES-ES-1p0_',ensmember,'_Annual_global.nc')
    
    try(nc <- nc_open(paste0(fn)))
    #try(localtime <- ncvar_get(nc, 'time'))
    
    # This part compensates for the fact that sometimes years are missing
    #try(localyear <- floor(2100 + (localtime / ysec)))
    #try(ix <- which(cn%in%localyear))
    
    try(dat <- extractTimeseries(nc, variable))
    
    try(datmat[i, ] <- dat)
    nc_close(nc)
  }
  datmat
}

getStandardMemberSSP <- function(ensloc, variable, nts = 251, cn = 1850:2100){
  
  datmat <- matrix(NA, nrow = 1, ncol = nts)
  colnames(datmat) <- cn
  
  ensmember <- 'S3'
  #fn <- paste0(ensloc,ensmember,'/stats/','JULES-ES-1p0_',ensmember,'_Annual_global.nc')
  fn <- paste0(ensloc,'JULES-ES-1p0_',ensmember,'_Annual_global.nc')
  
  try(nc <- nc_open(paste0(fn)))
  try(dat <- extractTimeseries(nc, variable))
  
  datmat[1, ] <- dat
  nc_close(nc)
  datmat
  
}



```



```{r}


ensloc_ssp585_S3 <- '/data/users/hadaw/JULES_ES_PPE/u-cf932/'
ensloc_ssp585_RAD <- '/data/users/hadaw/JULES_ES_PPE/u-ck647/'

ysec = 60*60*24*365
sspyears <- 1850:2100



```


```{r}
y_names_select_ssp585 <-  c("npp", "nbp", "cSoil", "cVeg",
                     "lai_lnd_mean",
                     "rh_lnd_sum" , "fLuc_lnd_sum", "fHarvest_lnd_sum",  
                     "treeFrac_lnd_mean" , "baresoilFrac_lnd_mean",
                     "shrubFrac_lnd_mean", "c3PftFrac_lnd_mean",
                     "c4PftFrac_lnd_mean"   
)

select_units <- c('GtC/year', 'GtC/year', 'GtC', 'GtC', 'index', 'GtC/year','GtC/year', 'GtC/year', '%', '%', '%', '%', '%')
names(select_units) <- y_names_select_ssp585

# Histograms of end of century values
yvec_ssp585 <- paste0(y_names_select_ssp585, "_ens_ssp585_S3")

yvec_anom_ssp585 <- paste0(y_names_select_ssp585, "_ens_anom_ssp585_S3")

```

[1] "baresoilFrac_lnd_mean_ens_ssp585_S3" "c3PftFrac_lnd_mean_ens_ssp585_S3"    "c4PftFrac_lnd_mean_ens_ssp585_S3"   
 [4] "cSoil_ens_ssp585_S3"                 "cVeg_ens_ssp585_S3"                  "fHarvest_lnd_sum_ens_ssp585_S3"     
 [7] "fLuc_lnd_sum_ens_ssp585_S3"          "lai_lnd_mean_ens_ssp585_S3"          "nbp_ens_ssp585_S3"                  
[10] "npp_ens_ssp585_S3"                   "rh_lnd_sum_ens_ssp585_S3"            "shrubFrac_lnd_mean_ens_ssp585_S3"   
[13] "treeFrac_lnd_mean_ens_ssp585_S3" 



```{r}

if (file.exists("ensemble_timeseries_ssp_2022-08-09.rdata")) {
  load("ensemble_timeseries_ssp_2022-08-09.rdata")
} else {
  
  # primary carbon cycle outputs
  npp_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "npp_nlim_lnd_sum", cn = sspyears) / (1e12/ysec)
  
  nbp_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "nbp_lnd_sum", cn = sspyears) / (1e12/ysec)
  
  cSoil_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "cSoil_lnd_sum", cn = sspyears) / 1e12
  
  cVeg_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "cVeg_lnd_sum", cn = sspyears) / 1e12
  
  lai_lnd_mean_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "lai_lnd_mean", cn = sspyears)
  
  # fluxes
  rh_lnd_sum_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "rh_lnd_sum", cn = sspyears) / (1e12/ysec)
  
  fLuc_lnd_sum_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "fLuc_lnd_sum", cn = sspyears) / (1e12/ysec)
 
  fHarvest_lnd_sum_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "fHarvest_lnd_sum", cn = sspyears) / (1e12/ysec)
  
  # fractions
  treeFrac_lnd_mean_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "treeFrac_lnd_mean", cn = sspyears)

  shrubFrac_lnd_mean_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "shrubFrac_lnd_mean", cn = sspyears)  
  
  baresoilFrac_lnd_mean_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "baresoilFrac_lnd_mean", cn = sspyears)  
  
  c3PftFrac_lnd_mean_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "c3PftFrac_lnd_mean", cn = sspyears) 
  
  c4PftFrac_lnd_mean_ens_ssp585_S3 <- makeTimeseriesEnsembleSSP(ensloc = ensloc_ssp585_S3,
                                                 nstart = 499, nend = 999, variable = "c4PftFrac_lnd_mean", cn = sspyears)   
  
  save(npp_ens_ssp585_S3, nbp_ens_ssp585_S3, cSoil_ens_ssp585_S3, cVeg_ens_ssp585_S3, lai_lnd_mean_ens_ssp585_S3, rh_lnd_sum_ens_ssp585_S3, fLuc_lnd_sum_ens_ssp585_S3, fHarvest_lnd_sum_ens_ssp585_S3,
       treeFrac_lnd_mean_ens_ssp585_S3, shrubFrac_lnd_mean_ens_ssp585_S3, baresoilFrac_lnd_mean_ens_ssp585_S3,c3PftFrac_lnd_mean_ens_ssp585_S3, c4PftFrac_lnd_mean_ens_ssp585_S3,
       file = "ensemble_timeseries_ssp_2022-08-09.rdata" )
  
}


```


```{r, include=FALSE}

# npp_stan_ssp585_S3<- getStandardMember(ensloc = ensloc_wave00, variable = "npp_nlim_lnd_sum") / (1e12/ysec)
# nbp_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "nbp_lnd_sum") / (1e12/ysec)
# cSoil_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "cSoil_lnd_sum") / 1e12
# cVeg_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "cVeg_lnd_sum") / 1e12
# lai_lnd_mean_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "lai_lnd_mean")
# 
# 
# # fluxes
# rh_lnd_sum_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "rh_lnd_sum") / (1e12/ysec)
# fLuc_lnd_sum_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "fLuc_lnd_sum") / (1e12/ysec)
# fHarvest_lnd_sum_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "fHarvest_lnd_sum") / (1e12/ysec)
# 
# 
# # fractions
# treeFrac_lnd_mean_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "treeFrac_lnd_mean")
# shrubFrac_lnd_mean_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "shrubFrac_lnd_mean")
# baresoilFrac_lnd_mean_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "baresoilFrac_lnd_mean")
# c3PftFrac_lnd_mean_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "c3PftFrac_lnd_mean")
# c4PftFrac_lnd_mean_stan_ssp585_S3 <- getStandardMember(ensloc = ensloc_wave00, variable = "c4PftFrac_lnd_mean")



```


```{r}
## ----------------------------------------------------------------------
## Anomalize ensemble
## ----------------------------------------------------------------------

npp_ens_anom_ssp585_S3 <- anomalizeTSmatrix(npp_ens_ssp585_S3, 1:20)
nbp_ens_anom_ssp585_S3 <- anomalizeTSmatrix(nbp_ens_ssp585_S3, 1:20)
cSoil_ens_anom_ssp585_S3 <- anomalizeTSmatrix(cSoil_ens_ssp585_S3, 1:20)
cVeg_ens_anom_ssp585_S3 <- anomalizeTSmatrix(cVeg_ens_ssp585_S3, 1:20)

rh_lnd_sum_ens_anom_ssp585_S3 <- anomalizeTSmatrix(rh_lnd_sum_ens_ssp585_S3, 1:20)
fLuc_lnd_sum_ens_anom_ssp585_S3 <- anomalizeTSmatrix(fLuc_lnd_sum_ens_ssp585_S3, 1:20)
lai_lnd_mean_ens_anom_ssp585_S3 <- anomalizeTSmatrix(lai_lnd_mean_ens_ssp585_S3, 1:20) 

fHarvest_lnd_sum_ens_anom_ssp585_S3 <- anomalizeTSmatrix(fHarvest_lnd_sum_ens_ssp585_S3, 1:20)
treeFrac_lnd_mean_ens_anom_ssp585_S3 <- anomalizeTSmatrix(treeFrac_lnd_mean_ens_ssp585_S3, 1:20)
shrubFrac_lnd_mean_ens_anom_ssp585_S3 <- anomalizeTSmatrix(shrubFrac_lnd_mean_ens_ssp585_S3, 1:20)
baresoilFrac_lnd_mean_ens_anom_ssp585_S3 <- anomalizeTSmatrix(baresoilFrac_lnd_mean_ens_ssp585_S3, 1:20)
c3PftFrac_lnd_mean_ens_anom_ssp585_S3 <- anomalizeTSmatrix(c3PftFrac_lnd_mean_ens_ssp585_S3, 1:20)
c4PftFrac_lnd_mean_ens_anom_ssp585_S3 <- anomalizeTSmatrix(c4PftFrac_lnd_mean_ens_ssp585_S3, 1:20)

#total_land_carbon_anom <- anomalizeTSmatrix(total_land_carbon_ens, 1:20)


```

```{r, fig.width = 12, fig.height = 12}

transpval <- 30
linecol = 'black'

par(mfrow= c(3,5), las = 1, mar = c(4,4,1,1))

matplot(sspyears, t(npp_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'NPP', xlab = '', ylab = 'GtC')

matplot(sspyears,t(nbp_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'NBP', xlab = '', ylab = 'GtC')

matplot(sspyears,t(cSoil_ens_ssp585_S3), 
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'cSoil', xlab = '', ylab = 'GtC')
 
matplot(sspyears,t(cVeg_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval), 
        main = 'cVeg', xlab = '', ylab = 'GtC' )

matplot(sspyears,t(rh_lnd_sum_ens_ssp585_S3 ),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'RH', xlab = '', ylab = 'GtC')

matplot(sspyears,t(fLuc_lnd_sum_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'fLuc', xlab = '', ylab = 'GtC')

matplot(sspyears,t(lai_lnd_mean_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'LAI', xlab = '', ylab = 'GtC')

matplot(sspyears,t(fHarvest_lnd_sum_ens_ssp585_S3), 
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'fHarvest', xlab = '', ylab = 'GtC')

matplot(sspyears,t(treeFrac_lnd_mean_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'treeFrac', xlab = '', ylab = '%')

matplot(sspyears,t(shrubFrac_lnd_mean_ens_ssp585_S3), 
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'shrubFrac', xlab = '', ylab = '%')

matplot(sspyears,t(baresoilFrac_lnd_mean_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'baresoilFrac', xlab = '', ylab = '%')

matplot(sspyears,t(c3PftFrac_lnd_mean_ens_ssp585_S3), type = 'l', 
        lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'c3PftFrac', xlab = '', ylab = '%')

matplot(sspyears,t(c4PftFrac_lnd_mean_ens_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'c4PftFrac', xlab = '',ylab = '%')


```


```{r, fig.width = 12, fig.height = 12}

transpval <- 30
linecol = 'black'

par(mfrow= c(3,5), las = 1, mar = c(4,4,1,1))

matplot(sspyears, t(npp_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'NPP', xlab = '', ylab = 'GtC')

matplot(sspyears,t(nbp_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'NBP', xlab = '', ylab = 'GtC')

matplot(sspyears,t(cSoil_ens_anom_ssp585_S3), 
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'cSoil', xlab = '', ylab = 'GtC')
 
matplot(sspyears,t(cVeg_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval), 
        main = 'cVeg', xlab = '', ylab = 'GtC' )

matplot(sspyears,t(rh_lnd_sum_ens_anom_ssp585_S3 ),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'RH', xlab = '', ylab = 'GtC')

matplot(sspyears,t(fLuc_lnd_sum_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'fLuc', xlab = '', ylab = 'GtC')

matplot(sspyears,t(lai_lnd_mean_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'LAI', xlab = '', ylab = 'GtC')

matplot(sspyears,t(fHarvest_lnd_sum_ens_anom_ssp585_S3), 
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'fHarvest', xlab = '', ylab = 'GtC')

matplot(sspyears,t(treeFrac_lnd_mean_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'treeFrac', xlab = '', ylab = '%')

matplot(sspyears,t(shrubFrac_lnd_mean_ens_anom_ssp585_S3), 
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'shrubFrac', xlab = '', ylab = '%')

matplot(sspyears,t(baresoilFrac_lnd_mean_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'baresoilFrac', xlab = '', ylab = '%')

matplot(sspyears,t(c3PftFrac_lnd_mean_ens_anom_ssp585_S3), type = 'l', 
        lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'c3PftFrac', xlab = '', ylab = '%')

matplot(sspyears,t(c4PftFrac_lnd_mean_ens_anom_ssp585_S3),
        type = 'l', lty = 'solid', col = makeTransparent(linecol, transpval),
        main = 'c4PftFrac', xlab = '',ylab = '%')


```


## Which ensemble members match constraints?

Start with Level 2 (Andy's constraints in NPP, NBP, cSoil and cVeg)

```{r}

ix_mv <- which(sspyears %in% 1996:2015)

yvec_constraints <- c('npp_ens_ssp585_S3', 'nbp_ens_ssp585_S3', 'cSoil_ens_ssp585_S3', 'cVeg_ens_ssp585_S3')

# Matrix of constraint output modern values (1996 - 2015)
Y_const_ssp585_mv <- matrix(data = NA, nrow = nrow(npp_ens_ssp585_S3), ncol = length(yvec_constraints))
colnames(Y_const_ssp585_mv) <- yvec_constraints

for(i in 1:length(yvec_constraints)){
  
  Y_trunc <- get(yvec_constraints[i])[, ix_mv]
  Y_trunc_mean <- apply(Y_trunc, 1, mean, na.rm = TRUE)
  Y_const_ssp585_mv[,i] <- Y_trunc_mean
}

```


```{r}

createConstraintString <- function(yvec, mins, maxes){
  # This function constructs a logical expression as a string to be evaluated
  out <- 'which('
  
  for(i in 1:length(yvec)){
    
    if(i<length(yvec)){
      
    subconst <- paste0('Y_unif','[,', '"',yvec[i],'"',']', '>', mins[i], '&', 'Y_unif','[,', '"',yvec[i], '"',']' , '<', maxes[i], '&')
    }
    
    else{
     subconst <-  paste0('Y_unif','[,', '"',yvec[i],'"',']', '>', mins[i], '&', 'Y_unif','[,', '"',yvec[i], '"',']' , '<', maxes[i])
    }
   out <-  paste0(out, subconst)
  }
  
  out <- paste0(out, ')')
}


AW_mins <- c(35, 0, 750, 300)
AW_maxes <- c(80, 10000, 3000, 800)

Y_unif <- Y_const_ssp585_mv
ix_kept_AW <- eval(parse(text = createConstraintString(yvec=yvec_constraints, mins = AW_mins, maxes = AW_maxes)))



#eval(parse(text = paste0()))

inputConstraintSize <- function(Y_unif, yvec, mins, maxes){
  # Calculate the indices of Y_unif that are within the bounds set by mins and maxes
  
  
  ix_kept <- eval(parse(text = createConstraintString(yvec=yvec, mins = mins, maxes = maxes)))
  
  prop_kept <- length(ix_kept) / nrow(Y_unif)
  
  return(list(ix_kept = ix_kept, prop_kept = prop_kept))
  
}
  
```

### Timeseries of absolute vales

```{r, fig.width = 10, fig.height = 12}
lcol_wave01 <- makeTransparent('firebrick',  80)
lcol_wave01_level2 <- makeTransparent('gold',  80)
stancol = 'black'

linePlotMultiEns <- function(years, ens1, ens2, col1, col2, ylab, main, ylim = NULL){
  # Plot wave00 and wave01 timeseries on top of one another
  
  nt <- length(years) 
  if(is.null(ylim)){
    
  #ylim = range(c(ens1[,1], ens1[,nt], ens2[,1], ens2[ ,nt]))
  ylim = range(c(ens1,ens2))
  }
  
  else ylim <- ylim
  
  matplot(years, t(ens1), type = 'l', lty = 'solid',ylim = ylim, col = col1,
        ylab = ylab, main = main, xlab = '',
        bty = 'n', lwd = 1.5)
  matlines(years, t(ens2), col = col2, lty = 'solid', lwd = 1.5)
}
```


```{r, fig.width = 12, fig.height = 12}

par(mfrow= c(3,5), las = 1, mar = c(4,4,1,1))

for(i in 1:length(yvec_ssp585)){

linePlotMultiEns(years = sspyears, ens1 = get(yvec_ssp585[i]),
                 ens2 = get(yvec_ssp585[i])[ix_kept_AW, ],
                 col1 = lcol_wave01, col2 = lcol_wave01_level2,
                 ylab = select_units[i], main = yvec_ssp585[i])
                 
}

reset()

legend('bottomright', legend = c('wave01', 'wave01 level2'), lty = 'solid', lwd = 2, col = c(lcol_wave01, lcol_wave01_level2), inset = c(0.05, 0.15) )



```


### Anomaly timeseries

```{r, fig.width = 10, fig.height = 12}

par(mfrow= c(3,5), las = 1, mar = c(4,4,1,1))

for(i in 1:length(yvec_ssp585)){

linePlotMultiEns(years = sspyears, ens1 = get(yvec_anom_ssp585[i]),
                 ens2 = get(yvec_anom_ssp585[i])[ix_kept_AW, ],
                 col1 = lcol_wave01, col2 = lcol_wave01_level2,
                 ylab = select_units[i], main = yvec_ssp585[i])
                 
}

reset()

legend('bottomright', legend = c('wave01', 'wave01 level2'), lty = 'solid', lwd = 2, col = c(lcol_wave01, lcol_wave01_level2), inset = c(0.05, 0.15) )

```


## Include baresoil fraction data

```{r}

load('treefrac/global_frac_cci.RData')

frac_cci <- c(global_frac_cci[1] + global_frac_cci[2], global_frac_cci[3], global_frac_cci[4], global_frac_cci[5], global_frac_cci[8] )
#frac_mv <- c(standard_modern_value['treeFrac_lnd_mean'], standard_modern_value['c3PftFrac_lnd_mean'], #standard_modern_value['c4PftFrac_lnd_mean'], standard_modern_value['shrubFrac_lnd_mean'], standard_modern_value['baresoilFrac_lnd_mean']  )


fraclab = c('Trees', 'C3 grasses', 'C4 grasses', 'Shrubs', 'Bare soil')


# Half and double the LC CCI data to get bounds for constraining JULES ES-1.0
frac_cci_max <- frac_cci + (0.5*frac_cci)
frac_cci_min <- frac_cci - (0.5*frac_cci)

```


```{r}

# Modern value 
ix_mv <- which(sspyears %in% 1996:2015)

yvec_constraints <- c('npp_ens_ssp585_S3', 'nbp_ens_ssp585_S3', 'cSoil_ens_ssp585_S3', 'cVeg_ens_ssp585_S3','baresoilFrac_lnd_mean_ens_ssp585_S3')

# Matrix of constraint output modern values (1996 - 2015)
Y_const_ssp585_mv <- matrix(data = NA, nrow = nrow(npp_ens_ssp585_S3), ncol = length(yvec_constraints))
colnames(Y_const_ssp585_mv) <- yvec_constraints

for(i in 1:length(yvec_constraints)){
  
  Y_trunc <- get(yvec_constraints[i])[, ix_mv]
  Y_trunc_mean <- apply(Y_trunc, 1, mean, na.rm = TRUE)
  Y_const_ssp585_mv[,i] <- Y_trunc_mean
}

```

```{r}
const_mins <- c(35, 0, 750, 300, frac_cci_min['bare soil']*100)
const_maxes <- c(80, 10000, 3000, 800, frac_cci_max['bare soil']*100)

Y_unif <- Y_const_ssp585_mv
ix_kept_AWBS <- eval(parse(text = createConstraintString(yvec=yvec_constraints, mins = const_mins, maxes = const_maxes)))

```


```{r, fig.width = 10, fig.height = 12}
lcol_wave01 <- makeTransparent('tomato2',  120)
lcol_wave01_level2 <- makeTransparent('gold',  120)
lcol_wave01_level3<- makeTransparent('cadetblue',  200)
stancol = 'black'


```

```{r, fig.width = 10, fig.height = 12}

par(mfrow= c(3,5), las = 1, mar = c(4,4,1,1))

for(i in 1:length(yvec_ssp585)){

linePlotMultiEns(years = sspyears, ens1 = get(yvec_ssp585[i]),
                 ens2 = get(yvec_ssp585[i])[ix_kept_AWBS, ],
                 col1 = lcol_wave01, col2 = lcol_wave01_level3,
                 ylab = select_units[i], main = yvec_ssp585[i])
                 
}

reset()

legend('bottomright', legend = c('wave01', 'wave01 level3'), lty = 'solid', lwd = 2, col = c(lcol_wave01, lcol_wave01_level3), inset = c(0.05, 0.15) )
```
### Anomaly timeseries AW + baresoil
```{r, fig.width = 12, fig.height = 12}

par(mfrow= c(3,5), las = 1, mar = c(4,4,1,1))

for(i in 1:length(yvec_ssp585)){

linePlotMultiEns(years = sspyears, ens1 = get(yvec_anom_ssp585[i]),
                 ens2 = get(yvec_anom_ssp585[i])[ix_kept_AWBS, ],
                 col1 = lcol_wave01, col2 = lcol_wave01_level3,
                 ylab = select_units[i], main = y_names_select_ssp585[i])
                 
}

reset()

legend('bottomright', legend = c('wave01', 'wave01 level3'), lty = 'solid', lwd = 2, col = c(lcol_wave01, lcol_wave01_level3), inset = c(0.05, 0.15) )
```
## End-Of-Century values under various constraints

Starting with AW constraints (level 2) and then bare soil (level3)

```{r}

#End of Century
ix_eoc <- which(sspyears %in% 2081:2100)
Y_ssp585_eoc <- matrix(data = NA, nrow = nrow(npp_ens_ssp585_S3), ncol = length(yvec_ssp585))
colnames(Y_ssp585_eoc) <- yvec_ssp585

for(i in 1:length(yvec_ssp585)){
  
  Y_trunc <- get(yvec_ssp585[i])[, ix_eoc]
  Y_trunc_mean <- apply(Y_trunc, 1, mean, na.rm = TRUE)
  Y_ssp585_eoc[,i] <- Y_trunc_mean
}
```

### Histograms of end-of-century absolute values

```{r, fig.width = 12, fig.height = 8}


hcol_wave01 <- makeTransparent('firebrick', 150)
hcol_wave01_level2 <- makeTransparent('gold',255)
hcol_wave01_level3 <- makeTransparent('cadetblue',255)


par(mfrow = c(3,5), fg = 'grey')

for(i in 1:length(yvec_ssp585)){
  
  h1 <- hist(Y_ssp585_eoc[,i], main = y_names_select_ssp585[i], xlab = select_units[i], col = hcol_wave01, ylim = c(0,225))
  
  hist(Y_ssp585_eoc[ix_kept_AW,i], col = hcol_wave01_level2, add = TRUE, breaks = h1$breaks)
  hist(Y_ssp585_eoc[ix_kept_AWBS,i], col = hcol_wave01_level3, add = TRUE, breaks = h1$breaks)
}

reset()

legend('bottomright', legend = c('wave01',  'wave01 level2', 'wave01 level3'),  fill= c(hcol_wave01,hcol_wave01_level2, hcol_wave01_level3), inset = c(0.05, 0.15) )

```

```{r, fig.width = 12, fig.height = 8}

par(mfrow = c(3,5), fg = 'grey')

for(i in 1:length(yvec_ssp585)){
  
  h1 <- hist(Y_ssp585_eoc[ix_kept_AW,i], main = y_names_select_ssp585[i], xlab = select_units[i], col = hcol_wave01_level2, ylim = c(0,80))
  hist(Y_ssp585_eoc[ix_kept_AWBS,i], col = hcol_wave01_level3, add = TRUE, breaks = h1$breaks)
}

reset()

legend('bottomright', legend = c( 'wave01 level2', 'wave01 level3'),  fill= c(hcol_wave01_level2, hcol_wave01_level3), inset = c(0.05, 0.15) )


```

### Histograms of end-of-century changes (anomaly from preindustrial)

```{r}

#End of Century
ix_eoc <- which(sspyears %in% 2081:2100)
Y_ssp585_anom_eoc <- matrix(data = NA, nrow = nrow(npp_ens_anom_ssp585_S3), ncol = length(yvec_ssp585))
colnames(Y_ssp585_anom_eoc) <- yvec_anom_ssp585

for(i in 1:length(yvec_anom_ssp585)){
  
  Y_trunc <- get(yvec_anom_ssp585[i])[, ix_eoc]
  Y_trunc_mean <- apply(Y_trunc, 1, mean, na.rm = TRUE)
  Y_ssp585_anom_eoc[,i] <- Y_trunc_mean
}
```

```{r, fig.width = 12, fig.height = 8}


#hcol_wave01 <- makeTransparent('firebrick', 150)
#hcol_wave01_level2 <- makeTransparent('gold',255)
#hcol_wave01_level3 <- makeTransparent('cadetblue',255)

par(mfrow = c(3,5), fg = 'grey')

for(i in 1:length(yvec_anom_ssp585)){
  
  h1 <- hist(Y_ssp585_anom_eoc[,i], main = y_names_select_ssp585[i], xlab = select_units[i], col = hcol_wave01, ylim = c(0,225))
  
  hist(Y_ssp585_anom_eoc[ix_kept_AW,i], col = hcol_wave01_level2, add = TRUE, breaks = h1$breaks)
  hist(Y_ssp585_anom_eoc[ix_kept_AWBS,i], col = hcol_wave01_level3, add = TRUE, breaks = h1$breaks)
  
  abline(v = 0)
}

reset()

legend('bottomright', legend = c('wave01',  'wave01 level2', 'wave01 level3'),  fill= c(hcol_wave01,hcol_wave01_level2, hcol_wave01_level3), inset = c(0.05, 0.15) )

```
```{r, fig.width = 12, fig.height = 8}


#hcol_wave01 <- makeTransparent('firebrick', 150)
#hcol_wave01_level2 <- makeTransparent('gold',255)
#hcol_wave01_level3 <- makeTransparent('cadetblue',255)

par(mfrow = c(3,5), fg = 'grey')

for(i in 1:length(yvec_anom_ssp585)){
  
 h1 <-  hist(Y_ssp585_anom_eoc[ix_kept_AW,i],  main = y_names_select_ssp585[i], xlab = select_units[i],col = hcol_wave01_level2, ylim = c(0,80))
  hist(Y_ssp585_anom_eoc[ix_kept_AWBS,i], col = hcol_wave01_level3, add = TRUE, breaks = h1$breaks)
  
  abline(v = 0)
}

reset()

legend('bottomright', legend = c('wave01 level2', 'wave01 level3'),  fill= c(hcol_wave01_level2, hcol_wave01_level3), inset = c(0.05, 0.15) )

```

## Adding cumulative NBP as a constraint

```{r}
test <- matrix(1:9, nrow = 3, byrow = TRUE)
test

t(apply(test, 1, FUN = cumsum))
```

```{r}

cnbp_ens_ssp585_S3 <-  t(apply(nbp_ens_ssp585_S3, 1, FUN = cumsum))

```


```{r}

# Modern value 
# ix_mv <- which(sspyears %in% 1996:2015)
# 
# yvec_constraints <- c('npp_ens_ssp585_S3', 'nbp_ens_ssp585_S3', 'cSoil_ens_ssp585_S3', 'cVeg_ens_ssp585_S3','baresoilFrac_lnd_mean_ens_ssp585_S3')
# 
# # Matrix of constraint output modern values (1996 - 2015)
# Y_const_ssp585_mv <- matrix(data = NA, nrow = nrow(npp_ens_ssp585_S3), ncol = length(yvec_constraints))
# colnames(Y_const_ssp585_mv) <- yvec_constraints
# 
# for(i in 1:length(yvec_constraints)){
#   
#   Y_trunc <- get(yvec_constraints[i])[, ix_mv]
#   Y_trunc_mean <- apply(Y_trunc, 1, mean, na.rm = TRUE)
#   Y_const_ssp585_mv[,i] <- Y_trunc_mean
# }

```


```{r}
# Calculate Modern value CNBP

ix_2014 <- which(sspyears == 2014)

cnbp_ens_ssp585_S3_mv <- cnbp_ens_ssp585_S3[, ix_2014]

breaks <- seq(from = -100, to = 350, by = 25)
hist(cnbp_ens_ssp585_S3_mv, main = "cumulative NBP since 1850 at 2014", col = hcol_wave01,
     xlab = 'Cumulative NBP since 1850 (GtC)', breaks = breaks)
hist(cnbp_ens_ssp585_S3_mv[ix_kept_AW], add = TRUE, col = hcol_wave01_level2, breaks = breaks)
hist(cnbp_ens_ssp585_S3_mv[ix_kept_AWBS], add = TRUE, col = hcol_wave01_level3, breaks = breaks)
     
reset()

legend('topright', legend = c('wave01',  'wave01 level2', 'wave01 level3'),  fill= c(hcol_wave01,hcol_wave01_level2, hcol_wave01_level3), inset = c(0.2, 0.2) )

```


```{r}
const_mins <- c(35, 0, 750, 300, frac_cci_min['bare soil']*100)
const_maxes <- c(80, 10000, 3000, 800, frac_cci_max['bare soil']*100)

Y_unif <- Y_const_ssp585_mv
ix_kept_AWBS <- eval(parse(text = createConstraintString(yvec=yvec_constraints, mins = const_mins, maxes = const_maxes)))

```


```{r}

const_mins_level4 <- c(35, 0, 750, 300, frac_cci_min['bare soil']*100, -40)
const_maxes_level4 <- c(80, 10000, 3000, 800, frac_cci_max['bare soil']*100, 40)

Y_unif <- cbind(Y_const_ssp585_mv, cnbp_ens_ssp585_S3_mv)
level4_constraints <- c('npp_ens_ssp585_S3', 'nbp_ens_ssp585_S3', 'cSoil_ens_ssp585_S3', 'cVeg_ens_ssp585_S3', 'baresoilFrac_lnd_mean_ens_ssp585_S3', 'cnbp_ens_ssp585_S3_mv')
ix_kept_level4 <- eval(parse(text = createConstraintString(yvec=level4_constraints, mins = const_mins_level4, maxes = const_maxes_level4)))

```


```{r, fig.width = 7, fig.height = 10}

matplot(sspyears, t(cnbp_ens_ssp585_S3), type = 'l', lty = 'solid', col = makeTransparent(hcol_wave01, 100),
        ylab = 'GtC', main = 'Cumulative NBP', xlab = '',
        bty = 'n', lwd = 2)

matlines(sspyears, t(cnbp_ens_ssp585_S3[ix_kept_AW, ]), lty = 'solid', col = makeTransparent(hcol_wave01_level2, 200), lwd = 2)
matlines(sspyears, t(cnbp_ens_ssp585_S3[ix_kept_AWBS, ]),lty = 'solid', col = makeTransparent(hcol_wave01_level3, 200), lwd = 2)
matlines(sspyears, t(cnbp_ens_ssp585_S3[ix_kept_level4, ]),lty = 'solid', col = makeTransparent('black', 200), lwd = 2)

reset()

legend('bottomleft', legend = c('wave01',  'wave01 level2', 'wave01 level3', 'wave01 level4'),  col = c(hcol_wave01,hcol_wave01_level2, hcol_wave01_level3, 'black'), inset = c(0.2, 0.2), lty = 'solid' )

```


```{r}
(length(ix_kept_AW) / 500) * 100
(length(ix_kept_AWBS) /500) * 100
(length(ix_kept_level4) / 500) * 100 
```



