---
title: "Sensitivity Analysis Earth System configuration of JULES"
author: "Doug McNeall"
date: "09/08/2021"
output: 
    html_notebook:
        toc: true
        toc_float: true
        toc_depth: 3
        number_sections: true
---


## Preliminaries
Load libraries, functions and data.

```{r, echo = FALSE, message = FALSE, warning=FALSE, results = 'hide'}
# Load helper functions

knitr::opts_chunk$set(fig.path = "figs/", echo = FALSE, message = FALSE, warnings = FALSE)

# load helper functions, data and do preliminary processing of the ensemble.
source('JULES-ES-1p0-common.R')
```


```{r}
# Helper functions

sensvar = function(oaat_pred, n, d){
  # Calculate variance as a global sensitivity meansure
  out = rep(NA,d)
  for(i in 1:d){
    ix = seq(from = ((i*n) - (n-1)), to =  (i*n), by = 1)
    out[i] = var(oaat_pred$mean[ix])
  }
  out
}


twoStep_sens <- function(X, y, n=21, predtype = 'UK', nugget=NULL, nuggetEstim=FALSE, noiseVar=NULL, seed=NULL, trace=FALSE, maxit=100,
                        REPORT=10, factr=1e7, pgtol=0.0, parinit=NULL, popsize=100){
  # Sensitivity analysis with twoStep emulator. 
  # Calculates the variance of the output varied one at a time across each input.
  d = ncol(X)
  X_norm <- normalize(X)
  X_oaat <- oaat_design(X_norm, n, med = TRUE)
  colnames(X_oaat) = colnames(X)
  
  twoStep_em = twoStep_glmnet(X=X, y=y, nugget=nugget, nuggetEstim=nuggetEstim, noiseVar=noiseVar,
                              seed=seed, trace=trace, maxit=maxit,
                              REPORT=REPORT, factr=factr, pgtol=pgtol,
                              parinit=parinit, popsize=popsize)
  
  oaat_pred = predict(twoStep_em$emulator, newdata = X_oaat, type = predtype)
  
  sens = sensvar(oaat_pred = oaat_pred, n=n, d=d)
  out = sens
  out
}



oaatSensvarKm <- function(X, y, n = 21, formula = ~., predtype = 'UK', ...){
  # one-at-a-time sensitivity summary with a standard dicekriging emulator
  
  d = ncol(X)
  X_norm <- normalize(X)
  X_oaat <- oaat_design(X_norm, n, med = TRUE)
  colnames(X_oaat) = colnames(X)
  

  em <- km(formula = formula, design = X, response = y, ...)
  
  oaat_pred = predict(em, newdata = X_oaat, type = predtype)
  
  sens = sensvar(oaat_pred = oaat_pred, n=n, d=d)
  out = sens
  out
  
}

```

```{r}
# We're just interested in the "sum" (global totals) data, not the "mean" (global means) data
y_names_sum <- c('nbp_lnd_sum', 'fLuc_lnd_sum', 'npp_nlim_lnd_sum', 'cSoil_lnd_sum',
            'cVeg_lnd_sum', 'landCoverFrac_lnd_sum', 'fHarvest_lnd_sum',
            'lai_lnd_sum', 'rh_lnd_sum', 'treeFrac_lnd_sum', 'c3PftFrac_lnd_sum', 
            'c4PftFrac_lnd_sum', 'shrubFrac_lnd_sum', 'baresoilFrac_lnd_sum')
```


```{r, warning=FALSE, message = FALSE}

if (file.exists("oat_twostep.rdata")) {
  load("oat_twostep.rdata")
} else {
oat_test <- twoStep_sens(X = X_level1, y = y_level1)


save(oat_test, file = "oat_twostep.rdata")
}

```


```{r}

if (file.exists("oaat_Y.rdata")) {
  load("oaat_Y.rdata")
} else {
  
  
oat_var_sensmat_Y <- matrix(NA, nrow = length(y_names_sum), ncol = ncol(X_level1))

for(i in 1:length(y_names_sum)){
  
  yname <- y_names_sum[i]
  y <- Y_level1[, yname]
  oat <- twoStep_sens(X = X_level1, y = y)
  oat_var_sensmat_Y[i, ] <- oat
}

save(y_names_sum, oat_var_sensmat_Y, file = "oaat_Y.rdata")
}
```



## Sensitivity at level1 constraint (truncated f0_io)
bwl_io becomes very influential across a number of variables

```{r, fig.width= 10, fig.height = 6}
# normalize the sensitivity matrix
colnames(oat_var_sensmat_Y) <- colnames(X_level1)
rownames(oat_var_sensmat_Y) <- y_names_sum

#test <- normalize(t(oat.var.sensmat))

#image(test)
#par()
heatmap(oat_var_sensmat_Y, Rowv = NA, Colv = NA, mar = c(10,10), scale = 'row')
heatmap(oat_var_sensmat_Y, mar = c(10,10), scale = 'row')

```


```{r, fig.width =12, fig.height = 7}


normsens_Y <- normalize(t(oat_var_sensmat_Y))

par(mar = c(12,12,5,2))
image(normsens_Y, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance across level 1 ensemble', side = 3, adj = 0, line = 2, cex = 1.8)

```

# Sensitivity at level1a constraint (f0_io and b_wl_io truncated)
Define constraint level 1a as those members that run, and have F0_io <0.9 & b_wl_io > 0.15 (normalised).
```{r}

y_level1a <- Y_level1a[, 'npp_nlim_lnd_sum']

oaat_sens_npp <- oaatSensvarKm(X = X_level1a, y = y_level1a)

```

```{r, fig.width = 7, fig.height = 7, results = 'hide', warning=FALSE}
em_level1a <- km(~., design = X_level1a,  response = y_level1a)
```


 # Plot the regular km emulator.

```{r, fig.width = 7, fig.height = 7}
plot(em_level1a)


```


```{r}
if (file.exists("oaat_level1a_Y.rdata")) {
  load("oaat_level1a_Y.rdata")
} else {
  
oat_var_sensmat_level1a_Y <- matrix(NA, nrow = length(y_names_sum), ncol = ncol(X_level1a))

for(i in 1:length(y_names_sum)){
  
  yname <- y_names_sum[i]
  y <- Y_level1a[, yname]
  oat <- oaatSensvarKm(X = X_level1a, y = y)
  oat_var_sensmat_level1a_Y[i, ] <- oat
}

save(y_names_sum, oat_var_sensmat_level1a_Y, file = "oaat_level1a_Y.rdata")
}

```


```{r, fig.width = 12, fig.height = 8}
normsens_level1a_Y <- normalize(t(oat_var_sensmat_level1a_Y))

par(mar = c(12,12,5,2))
image(normsens_level1a_Y, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1a), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance across level 1a ensemble', side = 3, adj = 0, line = 2, cex = 1.8)


```

# Anomaly sensitivities

```{r}
if (file.exists("oaat_level1a_YAnom.rdata")) {
  load("oaat_level1a_YAnom.rdata")
} else {
  
oat_var_sensmat_level1a_YAnom <- matrix(NA, nrow = length(y_names_sum), ncol = ncol(X_level1a))

for(i in 1:length(y_names_sum)){
  
  yname <- y_names_sum[i]
  y <- YAnom_level1a[, yname]
  oat <- oaatSensvarKm(X = X_level1a, y = y)
  oat_var_sensmat_level1a_YAnom[i, ] <- oat
}

save(y_names_sum, oat_var_sensmat_level1a_YAnom, file = "oaat_level1a_YAnom.rdata")
}

```



```{r, fig.width = 12, fig.height = 7}



normsens_level1a_YAnom <- normalize(t(oat_var_sensmat_level1a_YAnom))

par(mar = c(12,12,5,2))
image(normsens_level1a_YAnom, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1a), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance across level 1a ensemble', side = 3, adj = 0, line = 2, cex = 1.8)


```

```{r, fig.width = 12, fig.height=6}
# Summarise the sensitivity summaries per parameter

normsens_mean_Y <- apply(normsens_level1a_Y ,1, mean)
normsens_mean_YAnom <- apply(normsens_level1a_YAnom ,1, mean)


rn <- range(normsens_mean_Y, normsens_mean_YAnom)
plot(normsens_mean_Y)
points(normsens_mean_YAnom, col = 'red')

plot(normsens_mean_Y,  normsens_mean_YAnom)
abline(0,1)



```


```{r, fig.height = 8, fig.width = 4}

sort_ix <- sort(normsens_mean_Y, decreasing = FALSE, index.return = TRUE)

dotchart(normsens_mean_Y[sort_ix$ix], labels = colnames(X_level1a)[sort_ix$ix], bg = zissou5[5] , xlim = c(0,0.8), pch = 19)
points(normsens_mean_YAnom[sort_ix$ix], 1:32, col = zissou5[1], pch = 19)
         
dotchart(normsens_mean_Y, labels = colnames(X_level1a), bg = zissou5[5] , xlim = c(0,0.8), pch = 19)
points(normsens_mean_YAnom, 1:32, col = zissou5[1], pch = 19)


#dotchart(data$sold, pch = 21, labels = data$month, bg = "green",
#         pt.cex = 1.5, xlim = range(data$expected, data$sold) + c(-2, 2))
#points(data$expected, 1:nrow(data), col = "red", pch = 19, cex = 1.5)

```


```{r, fig.width = 12, fig.height = 12}

p <- ncol(X_level1a)

y_level1a <- Y_level1a[,'npp_nlim_lnd_sum']

par(mfrow = c(5,7), mar = c(3,1,3,1))
for(i in 1:p){
  plot(X_level1a[,i], y_level1a, xlab = '', ylab = '', main = colnames(X_level1a)[i], xlim = c(0,1))
}

```

# comparision of sensitivities level1 vs level1a
With b_wl truncated, we can clearly see the sensitivities change.

```{r, fig.width= 10, fig.height = 6}
# normalize the sensitivity matrix
colnames(oat_var_sensmat_level1a_Y) <- colnames(X_level1a)
rownames(oat_var_sensmat_level1a_Y) <- y_names_sum

#test <- normalize(t(oat.var.sensmat))

#image(test)
#par()
heatmap(oat_var_sensmat_level1a_Y, Rowv = NA, Colv = NA, mar = c(10,10), scale = 'row')
heatmap(oat_var_sensmat_level1a_Y, mar = c(10,10), scale = 'row')

```


```{r, fig.width =12, fig.height = 7}

normsens_level1a_Y <- normalize(t(oat_var_sensmat_level1a_Y))

par(mar = c(12,12,5,2))
image(normsens_level1a_Y, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1a), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance across level 1a ensemble', side = 3, adj = 0, line = 2, cex = 1.8)

```



```{r}
# This fails for "cVeg_lnd_sum"
# The result will be that some of the columns are repeated.

if (file.exists("oaat_YAnom.rdata")) {
  load("oaat_YAnom.rdata")
} else {

oat_var_sensmat_YAnom <- matrix(NA, nrow = length(y_names_sum), ncol = ncol(X_level1))

for(i in 1:length(y_names_sum)){
  
  yname <- y_names_sum[i]
  y <- YAnom_level1[, yname]
  try(oat <- twoStep_sens(X = X_level1, y = y))
  oat_var_sensmat_YAnom[i, ] <- oat
}

save(y_names_sum, oat_var_sensmat_YAnom, file = "oaat_YAnom.rdata")

}
```



```{r, fig.width= 10, fig.height = 6}
# normalize the sensitivity matrix
colnames(oat_var_sensmat_YAnom) <- colnames(X_level1)
rownames(oat_var_sensmat_YAnom) <- y_names_sum

#test <- normalize(t(oat.var.sensmat))

#image(test)
#par()
heatmap(oat_var_sensmat_YAnom, Rowv = NA, Colv = NA, mar = c(10,10), scale = 'row')
heatmap(oat_var_sensmat_YAnom, mar = c(10,6), scale = 'row')

```


```{r, fig.width =12, fig.height = 7}

normsens_YAnom <- normalize(t(oat_var_sensmat_YAnom))

par(mar = c(12,12,5,2))
image(normsens_YAnom, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance of historical change across level 1 ensemble', side = 3, adj = 0, line = 2, cex = 1.8)

```


# Plot both sensitivity matrices on top of one another.
It looks as though both the absolute value and the change over time are controlled by the same parameters.
```{r, fig.width =12, fig.height = 12}
par(mfrow = c(2,1), mar = c(12,12,5,2))

image(normsens_Y, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance of historical change across level 1 ensemble', side = 3, adj = 0, line = 2, cex = 1.8)


image(normsens_YAnom, axes = FALSE, col = blues)
axis(1, at = seq(from = 0, to = 1, length.out = p), labels = colnames(X_level1), las = 2)
axis(2, at = seq(from = 0, to = 1, length.out = length(y_names_sum)), labels = y_names_sum, las = 1)
mtext('One-at-a-time sensitivity, variance of historical change across level 1 ensemble', side = 3, adj = 0, line = 2, cex = 1.8)


```


Need Y and Yanom sensitivities at level 1a

```{r}
# This fails for "cVeg_lnd_sum"
# The result will be that some of the columns are repeated.

if (file.exists("oaat_level1a_YAnom.rdata")) {
  load("oaat_level1a_YAnom.rdata")
} else {

oat_var_sensmat_level1a_YAnom <- matrix(NA, nrow = length(y_names_sum), ncol = ncol(X_level1a))

for(i in 1:length(y_names_sum)){
  
  yname <- y_names_sum[i]
  y <- YAnom_level1a[, yname]
  try(oat <- twoStep_sens(X = X_level1a, y = y))
  oat_var_sensmat_level1a_YAnom[i, ] <- oat
}

save(y_names_sum, oat_var_sensmat_level1a_YAnom, file = "oaat_level1a_YAnom.rdata")

}
```



## One-at-a-time sensitivity analysis of constraining variables for understanding model response
"Constraining variables" being those we use to constrain the model (npp, nbp, cSoil and cVeg).
It's hard to maintain a high vegetation carbon in particular.  

Further idea: What parameter values might you choose to do this, and what might be the trade-offs you have to make?



```{r}
fit_list_const_level1a <- createKmFitList(X = X_level1a, Y = Y_const_level1a_scaled)

```


```{r}
# Check that oatSensVar and the plotting make sense


oaat_sens_cVeg <- oaatSensvarKm(X = X_level1a, y = Y_const_level1a_scaled[,"cVeg_lnd_sum"])

X_oaat_level1a <- oaat_design(X_level1a, n=21, med = TRUE)

colnames(X_oaat_level1a) = colnames(X)

y_oaat <- predict.km(fit_list_const_level1a[[4]], newdata = X_oaat_level1a, type = 'UK')

```


First, what parameters affect vegetation carbon and how? How sure are we about that?

```{r, fig.width = 8, fig.height = 10}


oaatLinePlot(X_oaat = X_oaat_level1a, y_oaat_mean = y_oaat$mean, y_oaat_sd = y_oaat$sd, 
             n_oaat = 21,nr = 6, nc = 6) 


```



```{r}

Y_oaat_const_level1a_scaled <- matrix(ncol = ncol(Y_const_level1a_scaled), nrow = nrow(X_oaat_level1a))

for(i in 1:ncol(Y_const_level1a_scaled)){

  y_oaat <- predict.km(fit_list_const_level1a[[i]], newdata = X_oaat_level1a, type = 'UK')
  Y_oaat_const_level1a_scaled[,i] <- y_oaat$mean
}

```


What might be the trade-offs for a high (or accurate) vegetation carbon? are they acceptable? Plot the oaat sensitivity of the other 3 outputs we're calibrating on. 

```{r, fig.width=10, fig.height = 10}
Y_oaat_const_level1a_scaled_norm <- normalize(Y_oaat_const_level1a_scaled)

oaatLinePlotMulti(X_oaat = X_oaat_level1a, Y_oaat = Y_oaat_const_level1a_scaled_norm ,  n_oaat = 21, nr = 6, nc = 6,
                  lwd = 3, cols = cbPal[c(1,2,6,8)])
  
reset()
legend('top', c('nbp', 'npp', 'csoil', 'cveg'), col = cbPal[c(1,2,6,8)], lty = 'solid', lwd = 3, horiz = TRUE)


```

