---
title: "Treefrac data"
output: html_notebook
---

```{r}
library(ncdf4)
library(ncdf4.helpers)
library(RColorBrewer)


yg <- brewer.pal(9, 'YlGn')
gyg <- c('grey80', yg[2:9])


```

```{r}

nc_cci <- nc_open('treefrac/qrparm.veg.frac_lc_cci_2005_9PFTs.nc')

```

```{r}

vars <- nc.get.variable.list(nc_cci)
print(vars)

```
```{r}
# This reads


lc_dat_cci <- ncvar_get(nc = nc_cci, varid = 'field1391' )


lat_cci <- ncvar_get(nc_cci,"latitude")
lon_cci <- ncvar_get(nc_cci,"longitude")


pseudo_cci <- ncvar_get(nc_cci,"pseudo")
  
```

The cci lats run -90 to 90 (145), whereas the land fraction runs -89.375 to 89.375 (144)

```{r, fig.width = 10, fig.height = 8}

lc_vec <- c('broadleaf trees', 'needleleaf trees', 'C3 (temperate) grass', 'C4 (tropical) grass', 'shrubs', 'urban', 'inland water', 'bare soil', 'ice')

```

This is the landfrac file from Andy - it has a smaller grid than the cci landcover file

```{r}

nc_landfrac <- nc_open('treefrac/bb075_m01s03i395.nc')

landfrac_dat <- ncvar_get(nc = nc_landfrac , varid = 'landFrac' )

image(landfrac_dat, col = gyg)

lat_landfrac <- ncvar_get(nc_landfrac,"latitude")
lon_landfrac <- ncvar_get(nc_landfrac,"longitude")

```

I created an area file using the following cdo command:

cdo gridarea qrparm.veg.frac_lc_cci_2005_9PFTs.nc gridarea.nc

Need to remap the landfrac file 
cdo remapbil,targetgrid ifile ofile

```{r}

nc_ga <- nc_open('treefrac/gridarea.nc')
vars <- nc.get.variable.list(nc_ga)
#print(vars)

ga_dat  <- ncvar_get(nc = nc_ga , varid = 'cell_area' )


lat_ga <- ncvar_get(nc_ga,"latitude")
lon_ga <- ncvar_get(nc_ga,"longitude")


#let's create a land mask from this map.

mask <- ga_dat

land_ix <- which(!is.na(lc_dat_cci[, , i]))
other_ix <- which(is.na(lc_dat_cci[, , i]))
mask[land_ix] <- 1
mask[other_ix] <- 0

image(mask, col = gyg)


```
```{r, fig.width = 10, fig.height = 9}

# It's going to be pretty close to just sum up with area * fraction

frac_area <- array(NA, dim = dim(lc_dat_cci))
  
  for(i in 1:dim(lc_dat_cci)[3]){
    
    frac_area[, , i] <- lc_dat_cci[,,i] * ga_dat
    
  }

frac_area_scaled <- frac_area / sum(ga_dat)

#pdf(file = 'figs/LC_CCI_maps.pdf', width = 10, height = 9)
par(mfrow = c(3,3), oma = c(8,0.1,0.1,0.1), las = 1)

for(i in 1:9){
  
image(lon_cci, lat_cci, lc_dat_cci[, , i], 
      main = lc_vec[i],
      xlab = 'longitude', ylab = 'latitude',
      zlim = c(0,1), col = gyg)
  
}


reset()

image.plot(lc_dat_cci[, , i],legend.only = TRUE, horizontal = TRUE, col = gyg, legend.shrink = 0.4, legend.width = 1,legend.args=list( text="Fraction", cex=1, side=3, line=0.5))
#dev.off()


```

We calculate the global fraction data as:  

sum(grid area * cci fraction) / global land surface area  

and global land surface area is sum(grid area * land mask)  



```{r}


# Global land surface area is the sum of the gridbox areas, multiplied by the land mask. 
glsa <- sum(ga_dat * mask)


# global sum of BL trees, for example


sum(frac_area[,,1], na.rm = TRUE) / glsa
sum(frac_area[,,2], na.rm = TRUE) / glsa

global_frac_cci <- rep(NA,9)

for(i in 1:9){
  
  global_frac_cci[i] <- sum(frac_area[,,i], na.rm = TRUE) / glsa
  
}

treefrac_total <- sum(global_frac_cci[1:2])

names(global_frac_cci) <- lc_vec

print(global_frac_cci * 100)

save(global_frac_cci, file = 'treefrac/global_frac_cci.RData')


```

